Cluster Difference Imaging Photometric Survey (CDIPS) II. TOI$\,$837: A Validated Jupiter-Sized Planet in IC~2602

### 2020-06-26 paper/ms.tex
update paper

Finding and studying systems undergoing these evolutionary changes is a major
goal in contemporary exoplanet research.  To identify stars younger than say
1$\,$Gyr, a number of direct and indirect methods are available
\citep{soderblom_ages_2010}.  The traditional approach is to isochronally
age-date coeval groups of stars, hereafter ``clusters'' \citep[{\it
e.g.},][]{lada_embedded_2003,zuckerman_young_2004,krumholz_star_2019}.  Young
field stars can also be identified using age indicators such as stellar
rotation periods, the abundance of photospheric lithium, and chromospheric
diagnostics such as calcium emission and broadband UV emission.  Studies by for
instance \citet{david_discovery_2018} and G{.}~Zhou et al. (2020, submitted)
have combined these methods for individual field stars.  Many of the latter
methods were summarized by \citet{mamajek_improved_2008}, and have since been
calibrated by {\it e.g.},
\citet{barnes_color-period_2015,meibom_spin-down_2015,angus_calibrating_2015}
and \citet{curtis_tess_2019} for stellar rotation,
\citet{zerjal_chromospherically_2017} for chromospheric activity, and {\it
e.g.}, \citet{berger_identifying_2018} and \citet{zerjal_galah_2019} for
lithium abundances.

@@

Reliable ages reported for \cn\ range from 30--46$\,$Myr.  We have
collected ages reported over the years in Table~\ref{tab:ages}.  The
Li depletion boundary technique yields slightly older absolute ages
than isochrone fitting.  Rather than redetermine the age of the
cluster and add another line to the table, we simply adopt the
absolute age range for \tn\ of 30--46$\,$Myr.

@@

TESS photometric rotation period = 3.5days ish.
vsini = 17.48 $\pm$ 0.15 (CHIRON).
The two agree.

For comparison, gyrochrones would ``predict'' X.XX.

This implies sub-Pleiades age expected for an IC 2602 member.

Figure~\ref{fig:lithium_rotation} compares the stellar rotation period of \tn\ to
periods that have been measured in a number of well-studied open clusters.
\tn\ seems to be gyrochronologically coeval with the Pleiades sequence.

We note that this is still in the ``sub 500 Myr'' range, when classical
Skumanich spin-down is not expected to be happening (why is this the number?).

This agrees with recent empirical results from Rebull et al 2020 and perhaps also 2018(?):
they observed the Taurus foreground (age 30 Myr) and saw that it was 
roughly consistent with the Pleaides.
Upper Sco had a consistent mean, but a much larger scatter.

%NOTE: begs the question... is there a large population of ~30-50 Myr
%old stars for which we could measure the rotation periods, and
%clarify if the scatter of Upper Sco chills out?

@@

Rafael's co-added FEROS spectra gave a Li doublet equivalent width of
154 milliangstrom.

Lithium depletion for a star of TOI-837's spectral type happens
over timescales >100 Myr (Soderblom+2013). This is because the
convective envelopes are very shallow, so they don't cycle material
down to the >3e6 K core until much later.

Nonetheless, comparison of early-G field stars to e.g., 600 Myr old
Hyads has shown that the depletion does happen over gigayear timescales
(Berger+2018, fig 7).

Figure~\ref{fig:lithium_rotation} compares our measured \tn\ Li EW 
to stars in the field, and to young open cluster members.
The field star measurements were collected by
\citet{berger_identifying_2018}.
We have shown the points with lithium detections at ${\rm SNR}>3$.
TOI-837 is consistent with being younger than almost all CKS
planet-hosting field stars.

The young open cluster members were selected for the presence of
lithium, as described by \citet{randich_gaiaeso_2018}.
The TOI-837 Li EW measured here (150 milliangstrom) is consistent
with what's seen for stars of similar colors in <100 Myr old moving
groups.
(Elliott+2016, Fig 11 -- attached).  This is sensible, because
IC 2602 is supposed to be 50 Myr old.

\subsubsection{Stellar Parameters}

@@

% TODO: decide whether to include

@@

\software{
  \texttt{astrobase} \citep{bhatti_astrobase_2018},
  %\texttt{astroplan} \citep{astroplan2018},
	\texttt{AstroImageJ} \citep{collins_astroimagej_2017},
  \texttt{astropy} \citep{astropy_2018},
  \texttt{astroquery} \citep{astroquery_2018},
  %\texttt{BATMAN} \citep{kreidberg_batman_2015},
  \texttt{cdips-pipeline} \citep{bhatti_cdips-pipeline_2019}
  \texttt{corner} \citep{corner_2016},
  %\texttt{emcee} \citep{foreman-mackey_emcee_2013},
  \texttt{exoplanet} \citep{exoplanet:agol19}
  \texttt{exoplanet} \citep{exoplanet:exoplanet}, and its
  dependencies \citep{exoplanet:agol19, exoplanet:kipping13, exoplanet:luger18,
  	exoplanet:theano}.
  \texttt{IPython} \citep{perez_2007},
	\texttt{isochrones} \citep{morton_2015_isochrones},
	\texttt{lightkurve} \citep{lightkurve_2018},
  \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
  \texttt{MESA} \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
  \texttt{numpy} \citep{walt_numpy_2011}, 
  \texttt{pandas} \citep{mckinney-proc-scipy-2010},
  \texttt{pyGAM} \citep{serven_pygam_2018_1476122},
  \texttt{PyMC3} \citep{salvatier_2016_PyMC3},
  \texttt{radvel} \citep{fulton_radvel_2018},
  %\texttt{scikit-learn} \citep{scikit-learn},
  \texttt{scipy} \citep{jones_scipy_2001}.
  \texttt{tesscut} \citep{brasseur_astrocut_2019},
	\texttt{VESPA} \citep{morton_efficient_2012,vespa_2015},
  \texttt{webplotdigitzer} \citep{rohatgi_2019},
  \texttt{wotan} \citep{hippke_wotan_2019}.
}

>>>>

Finding and studying systems undergoing these evolutionary changes is
a major goal in contemporary exoplanet research.  To identify stars
younger than say 1$\,$Gyr, a number of direct and indirect methods are
available \citep{soderblom_ages_2010}.  The traditional approach is to
isochronally age-date coeval groups of stars, hereafter ``clusters''
\citep[{\it
e.g.},][]{lada_embedded_2003,zuckerman_young_2004,krumholz_star_2019}.
Young field stars can also be identified using age indicators such as
stellar rotation periods, the abundance of photospheric lithium, and
chromospheric diagnostics such as calcium emission and broadband UV
emission.  Studies by for instance \citet{david_discovery_2018} and
G{.}~Zhou et al. (2020, submitted) have combined these methods for
individual field stars.  Many of the latter methods were summarized by
\citet{mamajek_improved_2008}, and have since been calibrated by {\it
e.g.},
\citet{irwin_rotational_2009,barnes_color-period_2015,meibom_spin-down_2015,angus_calibrating_2015}
and \citet{curtis_tess_2019} for stellar rotation,
\citet{zerjal_chromospherically_2017} for chromospheric activity, and
{\it e.g.}, \citet{elliott_search_2016,berger_identifying_2018} and
\citet{zerjal_galah_2019} for lithium abundances.

@@

Reliable ages reported for \cn\ range from 30--46$\,$Myr.  We have
collected ages reported over the years in Table~\ref{tab:ages}.  The
Li depletion boundary technique yields slightly older absolute ages
than isochrone fitting \citep{dobbie_ic_2010,randich_gaiaeso_2018}.
Rather than redetermine the age of the cluster and add another line to
the table, we simply adopt the absolute age range for \tn\ of
30--46$\,$Myr.

@@

As stars get older, their rotation rates incrementally slow due to
magnetic braking \cite{weber_angular_1967,skumanich_time_1972}.  One
way to verify the youth of \tn\ is by comparing its rotation period to
other stars with known ages.

We measured the rotation period from the PDCSAP light curve using the
Lomb-Scargle periodogram implemented in \texttt{astropy}
\citep{lomb_1976,scargle_studies_1982,vanderplas_periodograms_2015}.
We fitted the light curve without masking out the transits or flares,
as these represent a small fraction of the overall time series.  To
derive the uncertainty on the best period, we fitted a gaussian to the
dominant peak, after first ensuring that we had oversampled the initial
frequency grid.  This gave a rotation period of $P_{\rm rot} = 2.987
\pm 0.056\,{\rm d}$ when allowing for a single Fourier terms in the
periodogram model, and $P_{\rm rot} = 3.004 \pm 0.053\,{\rm d}$ when
allowing for two Fourier terms.  As the second model provides a better
fit to the data, we adopt the second measurement.

As we will discuss in Section~\ref{subsec:starparams}, we measured the
star's radius by combining the spectroscopic effective temperature
with a broadband photometry SED fit.  We would expect, combining our
$R_\star$ and $P_{\rm rot}$ measurements, that the equatorial velocity
$v$ of the star would be $17.67 \pm 0.32
\,{\rm km\,s}^{-1}$.  Our spectroscopically measured $v\sin i$ from
CHIRON, $17.48 \pm 0.15 \,{\rm km\,s}^{-1}$ agrees reasonably well
with this expectation.

The star is clearly a rapid rotator.
Figure~\ref{fig:lithium_rotation} compares its rotation period with
rotation periods that have been measured in a number of well-studied
open clusters.  \tn\ seems to be gyrochronologically coeval with the
Pleiades sequence.  This is not to say that \tn\ is ``Pleaides-aged'',
because the observed scatter in the rotation-period diagram for the
first 10--100$\,$Myr is quite high \citep[see Figure~9 of][]{rebull_rotation_2020}.  Instead, we interpret it as a
piece of evidence to support the claim that \tn\ is younger than $\sim
500\,{\rm Myr}$.

@@

Lithium depletion for early G-dwarfs like \tn\ requires
hundreds of megayears \citep{soderblom_ages_2014}. This is because
their convective envelopes are shallow, and so transport of
photospheric lithium to the hot core takes place over diffusive timescales,
rather than convective timescales.  Nonetheless, comparison of early G-dwarfs in
the field to {\it e.g.}, 600 Myr old Hyads has shown that the
depletion does indeed happen over many gigayears
\citep{berger_identifying_2018}.

The spectra of \tn\ all show the 6708$\,$\AA\ lithium doublet in
absorption. Opting to use our FEROS spectra because of their high S/N,
we measured the line's equivalent width (EW) to be $154 \pm 9 \, $m\AA.  Figure~\ref{fig:lithium_rotation} compares this EW to stars
in the field, and other young open cluster members.  The field star
measurements were collected by \citet{berger_identifying_2018}; we
show their reported lithium detections with ${\rm S/N}>3$.  The young
open cluster members were selected for the presence of lithium, as
described by \citet{randich_gaiaeso_2018}.  The measured \tn\ Li EW is
much larger than observed for field stars, and is consistent with
lithium absorption seen in stars with similar colors in sub-100$\,{\rm Myr}$
moving groups.

\subsubsection{Stellar Parameters}
\label{subsec:starparams}

@@

some words

@@

\software{
  \texttt{astrobase} \citep{bhatti_astrobase_2018},
  %\texttt{astroplan} \citep{astroplan2018},
	\texttt{AstroImageJ} \citep{collins_astroimagej_2017},
  \texttt{astropy} \citep{astropy_2018},
  \texttt{astroquery} \citep{astroquery_2018},
  %\texttt{BATMAN} \citep{kreidberg_batman_2015},
  \texttt{cdips-pipeline} \citep{bhatti_cdips-pipeline_2019}
  \texttt{corner} \citep{corner_2016},
  %\texttt{emcee} \citep{foreman-mackey_emcee_2013},
  \texttt{exoplanet} \citep{exoplanet:agol19}
  \texttt{exoplanet} \citep{exoplanet:exoplanet}, and its
  dependencies \citep{exoplanet:agol19, exoplanet:kipping13, exoplanet:luger18,
  	exoplanet:theano}.
  \texttt{IPython} \citep{perez_2007},
	\texttt{isochrones} \citep{morton_2015_isochrones},
	\texttt{lightkurve} \citep{lightkurve_2018},
  \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
  \texttt{MESA} \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
  \texttt{numpy} \citep{walt_numpy_2011}, 
  \texttt{pandas} \citep{mckinney-proc-scipy-2010},
  \texttt{pyGAM} \citep{serven_pygam_2018_1476122},
  \texttt{PyMC3} \citep{salvatier_2016_PyMC3},
  \texttt{radvel} \citep{fulton_radvel_2018},
  %\texttt{scikit-learn} \citep{scikit-learn},
  \texttt{scipy} \citep{jones_scipy_2001},
  \texttt{tesscut} \citep{brasseur_astrocut_2019},
	\texttt{VESPA} \citep{morton_efficient_2012,vespa_2015},
  \texttt{webplotdigitzer} \citep{rohatgi_2019},
  \texttt{wotan} \citep{hippke_wotan_2019}.
}
